---
output:
  pdf_document: default
  html_document: default
---
```{r}
library(dplyr)
library(grf) # for causal forest
```

Dataset prepared needs to be imported
```{r}
data_sum_click<-read.csv("data_vle_sumclicks.csv", stringsAsFactors = FALSE)

```

Designation of levels for categorical variables in the dataset
```{r}
data_sum_click$code_module<-factor(data_sum_click$code_module,levels = c("AAA","BBB", "CCC",
                                                                         "DDD", "EEE", "FFF",
                                                                         "GGG"))
data_sum_click$code_presentation<-factor(data_sum_click$code_presentation,levels =
                                           c('2013B','2013J', '2014B' ,'2014J'))
data_sum_click$imd_band<-factor(data_sum_click$imd_band,levels = 
                                  c('0-10%', '10-20%', '20-30%', '30-40%', '40-50%', '50-60%',
                                    '60-70%', '70-80%', '80-90%', '90-100%'))
data_sum_click$gender<-factor(data_sum_click$gender,levels = c('M','F'))
data_sum_click$region<-factor(data_sum_click$region,levels = c('East Anglian Region', 
                'Scotland', 'North Western Region',  'South East Region', 'West Midlands Region', 
                 'Wales', 'North Region','South Region', 'Ireland', 'South West Region', 
                'East Midlands Region', 'Yorkshire Region', 'London Region'))
data_sum_click$highest_education<-factor(data_sum_click$highest_education,levels =
                                           c('No Formal quals','Lower Than A Level', 
                                             'A Level or Equivalent','HE Qualification',
                                             'Post Graduate Qualification'))
data_sum_click$age_band<-factor(data_sum_click$age_band,levels = c('0-35','35-55','55-'))
data_sum_click$disability<-factor(data_sum_click$disability,levels = c('N','Y'))
data_sum_click$final_result<-factor(data_sum_click$final_result,levels =
                                      c('Fail','Withdrawn','Pass',"Distinction"))
```

Conversion of final result to binary variable, outcome of 1 denotes success of student (Pass or Distinction)
```{r}
data_sum_click<-data_sum_click%>%
  dplyr::mutate(result = ifelse(final_result=="Fail",0,
                         ifelse(final_result=="Withdrawn",0,1)))
data_sum_click<-replace(data_sum_click,is.na(data_sum_click),0)
```

Linear probability model to test which drivers are having a significant impact
Observations
1. All attributes except Code_presentation are statistically significant
2. Highest_education level, imd_band, and interactions have highest estimate and thus highest impact on probability of success of students.

```{r}
fit1<-lm(result~code_module+code_presentation+gender+region+highest_education+imd_band+age_band+
           num_of_prev_attempts+studied_credits+disability+dataplus+dualpane+externalquiz+
           folder+forumng+ glossary+homepage+htmlactivity+oucollaborate+ oucontent+ 
           ouelluminate+ ouwiki + page + questionnaire + quiz+ repeatactivity + 
           resource + sharedsubpage + subpage + url,data = data_sum_click)
summary(fit1)
```

Linear probability model including only drivers without interactions with VLE for comparison. VLE interactions show high predictive value and are recommended to be included based on causal analysis to get better estimates
```{r}
fit2<-lm(result~code_module+code_presentation+gender+region+highest_education+imd_band+age_band+num_of_prev_attempts+studied_credits+disability,data = data_sum_click)
summary(fit2)
```
Creation and Analysis of Causal Forests
Causal forests are used to implement causal analysis and capture complex relationships here. VLE interaction terms are included and Average treatment effect(ATE) is calculated for each attribute. Causal forest use a splitting criteria to maximise difference in relationship between treatment and outcome variables in between splits. However, making a causal tree does not validate a causal relationship.
ATE is calculated for each of the 41 columns generated with different levels of available attributes.
Observations:
1. Code module has the highest impact. This highlights the need to have uniform marking policies what may be able to remove any benefit that students may get with selection of courses
2. Highest education, disability, and IMD band are seen as the most important factors affecting student success. This may highlight need to invest in accessibility which may assist students with disability. IMD band shows the need for investment to provide additional assistance to students from poorer financial backgrounds
3. Gender, age, and studied credits have smaller impact on student performance

```{r}


split <- sample(c(FALSE, TRUE), nrow(data_sum_click), replace = TRUE, prob = c(0.2,0.8))
df.train <- data_sum_click[split,]
df.hold <- data_sum_click[!split,]


treatment<-df.train%>%
  dplyr::select(code_module,code_presentation,gender,region,highest_education,imd_band,age_band, disability,studied_credits,num_of_prev_attempts)%>%
  dplyr::mutate(gender_fac = ifelse(gender=='M',0,1))

treatment <- as.matrix(treatment$gender_fac)

# Isolate the outcome as a matrix
outcome <- as.matrix(df.train$result)


```

```{r}

X <- model.matrix(lm(outcome ~ -1 + dataplus+ dualpane+ externalquiz+ folder+ forumng+ glossary+ homepage+ htmlactivity+ oucollaborate+ oucontent +ouelluminate +ouwiki +page +questionnaire +quiz + repeatactivity +resource +sharedsubpage +subpage +url , data = df.train))

treatment.matrix <-model.matrix(lm(outcome ~ -1 + factor(code_module) + factor(code_presentation)  + factor(region) + factor(highest_education) + factor(imd_band) + factor(age_band) + factor(disability) + studied_credits + num_of_prev_attempts + factor(gender), data = df.train))

cfall<-list()
importance_all<-data.frame(cbind(colnames(X)))
ate<-data.frame()
for (i in seq(1,ncol(treatment.matrix))){
  treatment<-treatment.matrix[,i]
  cf <- causal_forest(X,outcome,treatment)
  ate<-rbind(ate,average_treatment_effect(cf, target.sample = 'overlap'))
  print(i)
  cfall<-c(cfall,list(cf))
  importance_all = (cbind(importance_all,variable_importance(cf)))
}



```

Output of files from model results
```{r}

ate = cbind(colnames(treatment.matrix),ate)
colnames(ate)<-c("Attribute","estimate","se")
ate$estimate_abs<-abs(ate$estimate)
ate <- ate %>% arrange(desc(estimate_abs))
ate
```

```{r}
write.csv(ate,"ate.csv")
write.csv(importance_all,"importance.csv")
```

